My wonderful paper

### 2024-12-06 _manuscript/_tex/index.tex
section 5

\caption{\label{fig-step-count}Number of days with fewer than 8,000
steps across 300 simulated 30-day periods. The orange bars indicate
instances where the count exeeds five days, representing an unexpected
outcome in this scenario.}

@@

\caption{\label{fig-logic-reg}Logic regression model fitted to the three
unit tests (test1, test2, test3) and the outcome expectation (unexpect)
as the response variable. The model suggests using an OR rule to combine
test1 and test3 to predict the outcome expectation.}

@@

A regression example is produced to illustrate the test selection
process for analyzing the effect of PM10 on mortality. The example
demonstrates how the procedure presented in Section~\ref{sec-method} can
be used to select cutoff values in the checks to diagnose an unexpected
PM10 coefficient from the generalized linear model.

Consider a generalized linear model (GLM) to study the effect of PM10 on
mortality {[}TODO: provide more context of the mortality-PM10 study{]}.
Analysts may expect a PM10 coefficient between {[}0, 0.005{]} after
considering the temperature confounding {[}TODO: reference{]}. This
expectation can be framed into a check that fails, labelled as 1, if the
PM10 coefficient is outside the range {[}0, 0.005{]}. Multiple factors
can impact the PM10 coefficient, such as the sample size, the strength
of the correlation between mortality and PM10, and the strength of the
correlation between mortality and temperature. Analysts may expect a
reasonable sample size to ensure the reliability of the coefficient
estimate. Outliers in the three variables can also leverage the
coefficient. While these are possible factors that could affect the
analysis result, it is not clear the cutoff values for these checks to
determine a failure. Here a list of checks are created in
Table~\ref{tbl-checks}.

@@

To generate replicate of the data, we first simulate the correlation
matrix of the three variables in a grid and then use a Gaussian copula
to generate a multivariate normal distribution based on the specified
correlation matrix and sample size. The multivariate normal distribution
is transformed using the normal CDF before the inverse CDF of the
assumed distributions of the three variables is applied. To determine
the appropriate distribution of each variable, various distributions are
fitted and compared. This includes poisson and negative binomial for
mortality; gamma, log-normal, exponential, weibull, and normal for pm10
and temperature; and beta for pm10 after rescaling the data to
\([0-1]\). AIC is used to determine the best distribution fit for each
variable with qq plot presented in Figure~\ref{fig-dist-fit} to evaluate
the fit. AIC suggests a negative binomial distributio nfor mortality, a
beta distribution for PM10 (multiple by 100 to obtain the original
scale), and a Weibull distribution for temperature. To include the
effect of outlier, we add a single outlier to the data for mortality and
PM10 {[}more details{]}.

@@

\caption{\label{fig-linear-reg-tree}Logic regression model fitted to the
twelve unit tests and the outcome expectation (unexpect) as the response
variable. The model suggests the relationship:
(\(\text{cor}(\text{m}, \text{tmp}) > -0.35\) AND
\(\text{smpl\_size} \le 600\)) OR
\(\text{cor}(\text{m},  \text{PM10}) < - 0.05\)}

@@

\caption{\label{tbl-linear-reg}Accuracy (precision and recall) and
parsimony (independence) metrics derived from the logic regression
model, along with harmonic and arithmetic means, for individual unit
tests (1: sample size, 2: mortality-PM10 correlation, 3:
mortality-temperature correlation), and the combined test rule 4:
(sample size AND mortality-temperature correlation) OR mortality-PM10
correlation.}

@@

A logic regression is fitted to predict whether the PM10 coefficient is
unexpected (outside the range of {[}0, 0.005{]}) using the checks listed
in Table~\ref{tbl-checks}. Precision, recall, and independence score,
along with their harmonic and arithmetic mean are calculated.
Figure~\ref{fig-linear-reg-tree} shows the logic regression tree from
the fitted model and Table~\ref{tbl-linear-reg} prints the numerical
summary of four selected single test and their combined test found by
the logic regression. The logic regression model picks up the following
cutoff value for each type of check:

@@

The tree structure suggests checking mortality-PM10 correlation and a
sample size larger than 200 with an additional check of either outlier
on mortality or correlation between mortality and temperature. This
combined check rule generates a 0.76 precision and a 0.88 recall for
predicting the unexpected PM10 coefficient. The single check,
\texttt{cor(m,\ PM10)\ \textless{}\ -0.03}, is also powerful with a high
precision of 0.988, but the low recall value of 0.61 suggests its high
false positive rate, as compared to the combined rule suggested by the
logic regression.

>>>>

\caption{\label{fig-step-count}Number of days with fewer than 8,000
steps across 300 simulated 30-day periods. The orange bars indicate
instances where the count exceeds five days, representing an unexpected
outcome in this scenario.}

@@

\caption{\label{fig-logic-reg}Logic regression model fitted to the three
unit tests (test1, test2, test3). The model suggests using an OR rule to
combine test1 and test3 to predict the outcome expectation.}

@@

Consider a generalized linear model (GLM) to study the effect of PM10 on
mortality {[}TODO: provide more context of the mortality-PM10 study{]}.
Analysts may expect a PM10 coefficient between {[}0, 0.005{]} after
considering the temperature confounding {[}TODO: reference{]}. The
example demonstrates how the procedure presented in
Section~\ref{sec-method} can be used to select cutoff values in the
checks to diagnose an unexpected PM10 coefficient from the generalized
linear model.

This PM10 coefficient expectation can be framed into a check that fails,
labelled as 1, if the PM10 coefficient is outside the range {[}0,
0.005{]} and 0 if it is within. Multiple factors can affect the PM10
coefficient, such as the sample size, the strength of the correlation
between mortality and PM10, and the strength of the correlation between
mortality and temperature. Analysts may expect a reasonable sample size
to ensure the reliability of the coefficient estimate. Outliers in the
three variables can also leverage the coefficient. While these are
possible factors that could affect the analysis result, it is not clear
the cutoff values for these checks to determine a failure. Here we
create a list of checks in Table~\ref{tbl-checks} with varied cutoff
values for each:

@@

To generate replicate of the data, we first simulate the correlation
matrix of the three variables in a grid and then use a Gaussian copula
to generate a multivariate normal distribution based on the specified
correlation matrix and sample size. The multivariate normal distribution
is transformed using the normal CDF before the inverse CDF of the
assumed distributions of the three variables is applied. To determine
the appropriate distribution of each variable, various distributions are
fitted and compared. This includes poisson and negative binomial for
mortality; gamma, log-normal, exponential, weibull, and normal for pm10
and temperature; and beta for pm10 after rescaling the data to
\([0-1]\). AIC is used to determine the best distribution fit for each
variable with qq plot presented in Figure~\ref{fig-dist-fit} to evaluate
the fit. AIC suggests a negative binomial distribution for mortality, a
beta distribution for PM10 (multiple by 100 to obtain the original
scale), and a Weibull distribution for temperature. To include the
effect of outlier, we add a single outlier to the data for both variable
mortality and PM10.

@@

\caption{\label{fig-linear-reg-tree}Logic regression model fitted to the
fourteen unit tests and the outcome expectation (unexpected) as the
response variable. The model suggests the relationship: (sample size
\emph{greater than} 200 AND mortality-PM10 correlation \textless{}
-0.04) AND (mortality-temperature correlation \textgreater{} -0.35 OR
there exist mortality outlier) to predict the unexpected PM10
coefficient. {[}sorry the proper plot is not ready{]}.}

@@

\caption{\label{tbl-linear-reg}Accuracy (precision and recall) and
parsimony (independence) metrics derived from the logic regression
model, along with harmonic and arithmetic means, for individual unit
tests (1: sample size, 2: mortality-PM10 correlation, 3:
mortality-temperature correlation, 4: mortality outlier), and the
combined test rule 5: (sample size AND mortality-PM10 correlation) AND
(mortality-temperature correlation OR mortality outlier.}

@@

A logic regression is fitted to predict whether the PM10 coefficient is
unexpected using the checks listed in Table~\ref{tbl-checks} and
Figure~\ref{fig-linear-reg-tree} shows the logic regression tree from
the fitted model. Precision, recall, and independence score, along with
their harmonic and arithmetic mean are calculated in
Table~\ref{tbl-linear-reg}. The logic regression model picks up the
following cutoff value for each type of check:

@@

The tree structure suggests checking mortality-PM10 correlation and a
sample size larger than 200 with an additional check of either outlier
on mortality or correlation between mortality and temperature. This
combined check rule generates a precision of 0.76 and a recall of 0.88
for predicting the unexpected PM10 coefficient. The single check,
\texttt{cor(m,\ PM10)\ \textless{}\ -0.03}, is also powerful with a high
precision of 0.988, but the low recall value of 0.61 suggests its high
false positive rate, as compared to the combined rule suggested by the
logic regression.

### 2024-12-06 index.qmd
section 5

```{r}
#| label: fig-step-count
#| fig-cap: Number of days with fewer than 8,000 steps across 300 simulated 30-day periods. The orange bars indicate instances where the count exeeds five days, representing an unexpected outcome in this scenario. 
set.seed(1234)
step_count <- tibble(id = 1:300) |> 
  rowwise() |> 
  mutate(
    small_n = rpois(1, 4),
    large_n = rpois(1, 4),
    norm_sd = rlogis(1, 800, scale = 100),
    norm_n = 30 - small_n - large_n,
    step = list(round(c(rnorm(small_n, 4000, 200), 
                        rnorm(large_n, 12000, 200), 
                        rnorm(norm_n, 9000, 300)
                        ), 0)),
    less_than_8000 = sum(step < 8000), 
    unexpect = ifelse(less_than_8000 > 5, 1, 0),
    test1 = ifelse(mean(step) < 8200, 1, 0),
    test2 = ifelse(quantile(step, 0.3, na.rm = TRUE) < 8300, 1, 0),
    test3 = ifelse(sd(step) > 2500, 1, 0),
    ) |>
  ungroup()

@@

```{r}
#| label: fig-logic-reg
#| fig.width: 6
#| fig.heigh: 2
#| fig-cap: Logic regression model fitted to the three unit tests (test1, test2, test3) and the outcome expectation (unexpect) as the response variable. The model suggests using an OR rule to combine test1 and test3 to predict the outcome expectation.
fit <- step_count |> fit_logic_reg(unexpect, test1:test3, seed = 1, nleaves = 4)
plot_logicreg(fit)
```

@@

A regression example is produced to illustrate the test selection process for analyzing the effect of PM10 on mortality. The example demonstrates how the procedure presented in @sec-method can be used to select cutoff values in the checks to diagnose an unexpected PM10 coefficient from the generalized linear model.

Consider a generalized linear model (GLM) to study the effect of PM10 on mortality [TODO: provide more context of the mortality-PM10 study]. Analysts may expect a PM10 coefficient between [0, 0.005] after considering the temperature confounding [TODO: reference]. This expectation can be framed into a check that fails, labelled as 1, if the PM10 coefficient is outside the range [0, 0.005]. Multiple factors can impact the PM10 coefficient, such as the sample size, the strength of the correlation between mortality and PM10, and the strength of the correlation between mortality and temperature. Analysts may expect a reasonable sample size to ensure the reliability of the coefficient estimate. Outliers in the three variables can also leverage the coefficient. While these are possible factors that could affect the analysis result, it is not clear the cutoff values for these checks to determine a failure. Here a list of checks are created in @tbl-checks.

@@

To generate replicate of the data, we first simulate the correlation matrix of the three variables in a grid and then use a Gaussian copula to generate a multivariate normal distribution based on the specified correlation matrix and sample size. The multivariate normal distribution is transformed using the normal CDF before the inverse CDF of the assumed distributions of the three variables is applied. To determine the appropriate distribution of each variable, various distributions are fitted and compared. This includes poisson and negative binomial for mortality; gamma, log-normal, exponential, weibull, and normal for pm10 and temperature; and beta for pm10 after rescaling the data to $[0-1]$. AIC is used to determine the best distribution fit for each variable with qq plot presented in @fig-dist-fit to evaluate the fit. AIC suggests a negative binomial distributio nfor mortality, a beta distribution for PM10 (multiple by 100 to obtain the original scale), and a Weibull distribution for temperature. To include the effect of outlier, we add a single outlier to the data for mortality and PM10 [more details].

@@

```{r}
#| fig.width: 8
#| fig.heigh: 3
fit <- pm10_df |> fit_logic_reg(unexpect, `smpl_size <= 200`:`mortality_outlier`, 
                                seed = 6, nleaves = 4)
plot_logicreg(fit) 
```
<!-- The caption with some LaTeX like this $\lambda$. -->

Logic regression model fitted to the twelve unit tests and the outcome expectation (unexpect) as the response variable. The model suggests the relationship:
($\text{cor}(\text{m}, \text{tmp}) > -0.35$ AND $\text{smpl\_size} \le 600$) OR $\text{cor}(\text{m},  \text{PM10}) < - 0.05$
:::

@@

tbl <- tibble(tests = c(1,2,3,4, 5)) |> bind_cols(df |> dplyr::select(-id)) 
tbl |> 
  kable(digits = 3, escape = FALSE,
        caption = "Accuracy (precision and recall) and parsimony (independence) metrics derived from the logic regression model, along with harmonic and arithmetic means, for individual unit tests (1: sample size, 2: mortality-PM10 correlation, 3: mortality-temperature correlation), and the combined test rule 4: (sample size AND mortality-temperature correlation) OR mortality-PM10 correlation.")

@@

A logic regression is fitted to predict whether the PM10 coefficient is unexpected (outside the range of [0, 0.005]) using the checks listed in @tbl-checks. Precision, recall, and independence score, along with their harmonic and arithmetic mean are calculated. @fig-linear-reg-tree shows the logic regression tree from the fitted model and @tbl-linear-reg prints the numerical summary of four selected single test and their combined test found by the logic regression. The logic regression model picks up the following cutoff value for each type of check:

@@

The tree structure suggests checking mortality-PM10 correlation and a sample size larger than 200 with an additional check of either outlier on mortality or correlation between mortality and temperature. This combined check rule generates a `r round(as.numeric(df[5, "precision"]), 3)` precision and a `r round(as.numeric(df[5, "recall"]), 3)` recall for predicting the unexpected PM10 coefficient. The single check, `cor(m, PM10) < -0.03`, is also powerful with a high precision of `r round(as.numeric(df[2, "precision"]), 3)`, but the low recall value of `r round(as.numeric(df[2, "recall"]), 3)` suggests its high false positive rate, as compared to the combined rule suggested by the logic regression.

>>>>

```{r}
#| label: fig-step-count
#| fig-cap: Number of days with fewer than 8,000 steps across 300 simulated 30-day periods. The orange bars indicate instances where the count exceeds five days, representing an unexpected outcome in this scenario. 
set.seed(1234)
step_count <- tibble(id = 1:300) |> 
  rowwise() |> 
  mutate(
    small_n = rpois(1, 4),
    large_n = rpois(1, 4),
    norm_sd = rlogis(1, 800, scale = 100),
    norm_n = 30 - small_n - large_n,
    step = list(round(c(rnorm(small_n, 4000, 200), 
                        rnorm(large_n, 12000, 200), 
                        rnorm(norm_n, 9000, 300)
                        ), 0)),
    less_than_8000 = sum(step < 8000), 
    unexpect = ifelse(less_than_8000 > 5, 1, 0),
    test1 = ifelse(mean(step) < 8200, 1, 0),
    test2 = ifelse(quantile(step, 0.3, na.rm = TRUE) < 8300, 1, 0),
    test3 = ifelse(sd(step) > 2500, 1, 0),
    ) |>
  ungroup()

@@

```{r}
#| label: fig-logic-reg
#| fig.width: 6
#| fig.heigh: 2
#| fig-cap: Logic regression model fitted to the three unit tests (test1, test2, test3). The model suggests using an OR rule to combine test1 and test3 to predict the outcome expectation.
fit <- step_count |> fit_logic_reg(unexpect, test1:test3, seed = 1, nleaves = 4)
plot_logicreg(fit)
```

@@

Consider a generalized linear model (GLM) to study the effect of PM10 on mortality [TODO: provide more context of the mortality-PM10 study]. Analysts may expect a PM10 coefficient between [0, 0.005] after considering the temperature confounding [TODO: reference]. The example demonstrates how the procedure presented in @sec-method can be used to select cutoff values in the checks to diagnose an unexpected PM10 coefficient from the generalized linear model.

This PM10 coefficient expectation can be framed into a check that fails, labelled as 1, if the PM10 coefficient is outside the range [0, 0.005] and 0 if it is within. Multiple factors can affect the PM10 coefficient, such as the sample size, the strength of the correlation between mortality and PM10, and the strength of the correlation between mortality and temperature. Analysts may expect a reasonable sample size to ensure the reliability of the coefficient estimate. Outliers in the three variables can also leverage the coefficient. While these are possible factors that could affect the analysis result, it is not clear the cutoff values for these checks to determine a failure. Here we create a list of checks in @tbl-checks with varied cutoff values for each:

@@

To generate replicate of the data, we first simulate the correlation matrix of the three variables in a grid and then use a Gaussian copula to generate a multivariate normal distribution based on the specified correlation matrix and sample size. The multivariate normal distribution is transformed using the normal CDF before the inverse CDF of the assumed distributions of the three variables is applied. To determine the appropriate distribution of each variable, various distributions are fitted and compared. This includes poisson and negative binomial for mortality; gamma, log-normal, exponential, weibull, and normal for pm10 and temperature; and beta for pm10 after rescaling the data to $[0-1]$. AIC is used to determine the best distribution fit for each variable with qq plot presented in @fig-dist-fit to evaluate the fit. AIC suggests a negative binomial distribution for mortality, a beta distribution for PM10 (multiple by 100 to obtain the original scale), and a Weibull distribution for temperature. To include the effect of outlier, we add a single outlier to the data for both variable mortality and PM10.

@@

```{r}
#| fig.width: 8
#| fig.heigh: 3
fit <- pm10_df |> fit_logic_reg(unexpect, `smpl_size <= 200`:`mortality_outlier`, 
                                seed = 6, nleaves = 4)
plot(fit) 
```

Logic regression model fitted to the fourteen unit tests and the outcome expectation (unexpected) as the response variable. The model suggests the relationship:
(sample size *greater than* 200 AND mortality-PM10 correlation < -0.04) AND (mortality-temperature correlation > -0.35 OR there exist mortality outlier) to predict the unexpected PM10 coefficient. [sorry the proper plot is not ready].
:::

@@

tbl <- tibble(tests = c(1,2,3,4, 5)) |> bind_cols(df |> dplyr::select(-id)) 
tbl |> 
  kable(digits = 3, escape = FALSE,
        caption = "Accuracy (precision and recall) and parsimony (independence) metrics derived from the logic regression model, along with harmonic and arithmetic means, for individual unit tests (1: sample size, 2: mortality-PM10 correlation, 3: mortality-temperature correlation,  4: mortality outlier), and the combined test rule 5: (sample size AND mortality-PM10 correlation) AND (mortality-temperature correlation OR mortality outlier.")

@@

A logic regression is fitted to predict whether the PM10 coefficient is unexpected using the checks listed in @tbl-checks and @fig-linear-reg-tree shows the logic regression tree from the fitted model. Precision, recall, and independence score, along with their harmonic and arithmetic mean are calculated in  @tbl-linear-reg. The logic regression model picks up the following cutoff value for each type of check:

@@

The tree structure suggests checking mortality-PM10 correlation and a sample size larger than 200 with an additional check of either outlier on mortality or correlation between mortality and temperature. This combined check rule generates a precision of `r round(as.numeric(df[5, "precision"]), 3)` and a recall of `r round(as.numeric(df[5, "recall"]), 3)` for predicting the unexpected PM10 coefficient. The single check, `cor(m, PM10) < -0.03`, is also powerful with a high precision of `r round(as.numeric(df[2, "precision"]), 3)`, but the low recall value of `r round(as.numeric(df[2, "recall"]), 3)` suggests its high false positive rate, as compared to the combined rule suggested by the logic regression.
